---
title: "Project 2"
author: "Eli (Ilya) Bolotin"
date: "10/21/2019"
output: word_document
fontsize: 11
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(statnet, devtools, UserNetR, igraphdata, maps, igraph)
```

# Load data
```{r}
data(USairports)
net1 <- simplify(USairports)
detach(package:igraph)
```

# Convert igraph to statnet network
```{r}
# df <- asDF(USairports) # this works
# net1 <- asNetwork(df$edges, vertices=df$vertexes)
df <- asDF(net1) # this works
net1 <- asNetwork(df$edges, vertices=df$vertexes)
```

# Analyze network
```{r}
# view vertex attributes
list.vertex.attributes(net1)

# view edge attributes
list.edge.attributes(net1)

# summary overall of network
# summary(net1)

# examine city vertex
head(get.vertex.attribute(net1, "City"))

# view density of network
density.n1 <- gden(net1)
```

Network density is `density.n1`. The closer to 1 the density is, "the more interconnected the network" [1] (p.14, class text). In the context of this data, density is small because of the hub-and-spoke nature of the airline industry. That is, the country consists of mostly regional airports that connect through nationally-central hubs for longer distance flights.

# Add state vertex

```{r}
net1 %v% "state" <- as.character(sapply(strsplit(get.vertex.attribute(net1, "City"), ", ", fixed = TRUE), function(x) x[2]))
```


# Examine complete network

```{r}
set.seed(30)
net2 <- symmetrize(net1)
gplot(net2, vertex.col = 2, displaylabels = FALSE)
```

The network is hard to intelligbly decipher due to 23k edges. But the pattern is clear: the network is divided into two primary communities that are connected by central nodes. 

# Subgraph the network

To get a better sense of the most central nodes, we create subgraphs based on 2 centrality measures: degree and betweenness. These measures are most relevant because they reflect the most connected nodes (degree) as well as how many sub-nodes are connected through the most connected nodes (betweenness).

```{r}
# add vertex attributes for each centrality measure
net1 %v% "degree" <- degree(net1)
net1 %v% "btw" <- betweenness(net1)

# view summary of centrality attributes
summary(get.vertex.attribute(net1, "degree"))
summary(get.vertex.attribute(net1, "btw"))

n1d <- get.inducedSubgraph(net1, which(net1 %v% "degree" >= 40))
n1b <- get.inducedSubgraph(net1, which(net1 %v% "btw" >= 2000))

# cities <- get.vertex.attribute(net1, "City")
# top_20 <- cities[which(net1 %v% "degree" >= 300)]
```

# plot subnetworks

```{r}
rescale <- function(nchar, low, high) {
    min_d <- min(nchar)
    max_d <- max(nchar)
    rscl <- ((high - low) * (nchar - min_d)) / (max_d - min_d) + low
    rscl
  }
```

```{r}
set.seed(30)

op <- par(mar = c(0,0,4,0),mfrow=c(1,2))
statelab1 = get.vertex.attribute(n1d, "state")
statelab2 = get.vertex.attribute(n1b, "state")

gplot(n1d, displaylabels = T, 
      label = statelab1, 
      label.cex = 0.7, 
      vertex.col = 2, 
      edge.col = "grey50", 
      edge.lwd = 0.05,
      displayisolates = F,
      usearrows = F)

gplot(n1b, displaylabels = T, 
      label = statelab2, 
      label.cex = 0.7, 
      vertex.col = 2, 
      edge.col = "grey50", 
      edge.lwd = 0.05,
      displayisolates = F,
      usearrows = F)
par(op)
```

```{r}
# delete.vertices(n1d, isolates(n1d))
# deg <- degree(n1d)
# gplot(n1d, displaylabels = F, label = statelab, vertex.cex = rescale(deg,1,6), vertex.col = 2, edge.col = "grey80", edge.lwd = 0.5)
```


# Visualization and Network Summary

# Community Detection

# Network modeling


